---
title: "Load_Data"
author: "Group 5"
date: "11/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```



```{r load-packages, echo=FALSE}
library(tidyverse)
library(extrafont)
library(mosaic)
library(vroom)
library(ggtext)
library(gapminder)
library(ggrepel)
library(patchwork)
library(gghighlight)
library(skimr)
library(sf)
library(lubridate)
library(janitor)
library(countrycode)
```


```{r}
esg_data_raw <- read_csv(here::here("data/ESG_csv","ESGData.csv"))
gdp_per_cap_raw <- read_csv(here::here("data","API_NY.GDP.PCAP.CD_DS2_en_csv_v2_3159040.csv"),skip = 3)


esg_data <- esg_data_raw %>% pivot_longer(cols=5:65, names_to = "Year") %>% clean_names() %>% select(-x2050,-x67) 
gdp_per_cap <- gdp_per_cap_raw %>% pivot_longer(cols=5:65, names_to = "Year") %>% clean_names() %>% select(-x66) 
esg_gdp_data <- rbind(esg_data, gdp_per_cap)

esg_data_final <- esg_gdp_data  %>%
  select(-indicator_code) %>%
  pivot_wider(names_from = indicator_name, values_from = value) %>% 
  mutate(continent = countrycode(sourcevar = country_name,
                            origin = "country.name",
                            destination = "continent")) %>% 
  clean_names()

write_csv(esg_data_final,"esg_data_final.csv")
colnames(esg_data_final)
esg_data_final$population_density_people_per_sq_km_of_land_area
```

Env:
 [3] "Adjusted savings: natural resources depletion (% of GNI)"                                                 
 [4] "Adjusted savings: net forest depletion (% of GNI)"                                                        
 [5] "Agricultural land (% of land area)"                                                                       
 [6] "Agriculture, forestry, and fishing, value added (% of GDP)"                                               
 [7] "Annual freshwater withdrawals, total (% of internal resources)"
 [11] "CO2 emissions (metric tons per capita)"  
 [13] "Cooling Degree Days (projected change in number of degree Celsius)"
 [14] "Droughts, floods, extreme temperatures (% of population, average 1990-2009)"   
 [16] "Electricity production from coal sources (% of total)"                                                    
[17] "Energy imports, net (% of energy use)"                                                                    
[18] "Energy intensity level of primary energy (MJ/$2011 PPP GDP)"                                              
[19] "Energy use (kg of oil equivalent per capita)"
[22] "Forest area (% of land area)"                                                                             
[23] "Fossil fuel energy consumption (% of total)"
[25] "GHG net emissions/removals by LUCF (Mt of CO2 equivalent)"
[29] "Heat Index 35 (projected change in days)"
[36] "Mammal species, threatened"                                                                               
[37] "Maximum 5-day Rainfall, 25-year Return Level (projected change in mm)"                                    
[38] "Mean Drought Index (projected change, unitless)"                                                          
[39] "Methane emissions (metric tons of CO2 equivalent per capita)" 
[42] "Nitrous oxide emissions (metric tons of CO2 equivalent per capita)" 
[44] "People using safely managed drinking water services (% of population)"
[45] "People using safely managed sanitation services (% of population)"
[46] "PM2.5 air pollution, mean annual exposure (micrograms per cubic meter)"
[56] "Renewable electricity output (% of total electricity output)"
[57] "Renewable energy consumption (% of total final energy consumption)"

Social:
  Livelihood:
 [1] "Access to clean fuels and technologies for cooking (% of population)"                                     
 [2] "Access to electricity (% of population)"
[10] "Children in employment, total (% of children ages 7-14)"
[34] "Life expectancy at birth, total (years)"
[50] "Poverty headcount ratio at national poverty lines (% of population)" 
[51] "Prevalence of overweight (% of adults)"                                                                   

  Poverty:
[21] "Food production index (2014-2016 = 100)"[26] "Gini index (World Bank estimate)"
[52] "Prevalence of undernourishment (% of population)"

  Births & Deaths:
[9] "Cause of death, by communicable diseases and maternal, prenatal and nutrition conditions (% of total)"
[20] "Fertility rate, total (births per woman)"
[30] "Hospital beds (per 1,000 people)"
[40] "Mortality rate, under-5 (per 1,000 live births)"
[66] "Unmet need for contraception (% of married women ages 15-49)"

  Education:
[35] "Literacy rate, adult total (% of people ages 15 and above)"  
[60] "School enrollment, primary (% gross)"                                                                     
[61] "School enrollment, primary and secondary (gross), gender parity index (GPI)"

  Equality:
[53] "Proportion of seats held by women in national parliaments (%)"
[54] "Ratio of female to male labor force participation rate (%) (modeled ILO estimate)"

  Economic:
 [8] "Annualized average growth rate in per capita real survey mean consumption or income, total population (%)"
[15] "Ease of doing business rank (1=most business-friendly regulations)"
[24] "GDP growth (annual %)"
[31] "Income share held by lowest 20%"                                                                          
[33] "Labor force participation rate, total (% of total population ages 15-64) (modeled ILO estimate)"          
[65] "Unemployment, total (% of total labor force) (modeled ILO estimate)" 

  Demographic:
[41] "Net migration"                                                                                          
[48] "Population ages 65 and above (% of total population)"                                                     
[49] "Population density (people per sq. km of land area)"                                                      

  Technology: 
[32] "Individuals using the Internet (% of population)"
[43] "Patent applications, residents"
[58] "Research and development expenditure (% of GDP)"
[62] "Scientific and technical journal articles"

Governance:

[12] "Control of Corruption: Estimate"                                                                          
[27] "Government Effectiveness: Estimate"                                                                       
[28] "Government expenditure on education, total (% of government expenditure)" 
[47] "Political Stability and Absence of Violence/Terrorism: Estimate"   
[55] "Regulatory Quality: Estimate"
[59] "Rule of Law: Estimate" 
[63] "Strength of legal rights index (0=weak to 12=strong)"  
[64] "Terrestrial and marine protected areas (% of total territorial area)"
[67] "Voice and Accountability: Estimate"  
