---
title: "ESG Group Project: Data cleaning & Visualizations"
author: "Group 5"
date: "`r Sys.Date()`"
output: 
    html_document:
      number_sections: no
      theme: cerulean
      toc: yes
      toc_depth: 3
      toc_float:
        collapsed: false
      fontsize: 10pt
      code_folding: show
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```



```{r load-packages, echo=FALSE}
library(tidyverse)
library(extrafont)
library(mosaic)
library(vroom)
library(ggtext)
library(janitor)
library(gapminder)
library(ggrepel)
library(patchwork)
library(gghighlight)
library(gridExtra)
library(skimr)
library(sf)
library(lubridate)
library(countrycode)

loadfonts(device="pdf")
# loadfonts(device="pdf") if you have a mac
```

## Data Cleaning

```{r data_cleaning}

esg_data_raw <- read_csv(here::here("data/ESG_csv","ESGData.csv"))

glimpse(esg_data_raw)
skimr::skim(esg_data_raw)
ncol(esg_data_raw) #67 (col 66 is year 2050 and col 67 is a variable with no information names X67)
#cols 5-65 contain yearly data from 1960-2020
#most indicators have missing data for certainyears. 

#long data 
esg_data <- esg_data_raw %>% pivot_longer(cols=5:65, names_to = "Year") %>% select(-`2050`,-`X67`) %>% janitor::clean_names()


## ALL OF THIS IS NOT SUPER USEFUL FOR ANALYIS (but still look)
# Information on sources of different country series
# esg_countryseries <- read_csv(here::here("data/ESG_csv","ESGCountry-Series.csv"))

#Information on country statistics and methodology
# esg_country <- read_csv(here::here("data/ESG_csv","ESGCountry.csv"))

 #Shows sources of data
# esg_footnote <- read_csv(here::here("data/ESG_csv","ESGFootNote.csv"))

#Explain meaning of different features related to time series
# esg_series_time <- read_csv(here::here("data/ESG_csv","ESGSeries-Time.csv"))
# esg_series <- read_csv(here::here("data/ESG_csv","ESGSeries.csv"))

```   


```{r}

#here we create the wide useable data with each variable as its own column
esg_data_final <- esg_data  %>% select(-indicator_code) %>% group_by(indicator_name) %>% pivot_wider(names_from = indicator_name, values_from = value) %>% janitor::clean_names()

#Writes to data folder
write.csv(esg_data_final, here::here("data","esg_data_final.csv"))

skim(esg_data_final)

esg_data_final$year <- as.integer(esg_data_final$year)


esg_data_final$continent = countrycode(sourcevar = esg_data_final$country_name,origin = "country.name",destination = "continent")

esg_data_final %>% count(continent) %>% mutate(perc = round(100*n/sum(n),2))

esg_data_final %>% count(year) %>% mutate(perc = round(100*n/sum(n),2))


```

## Animation of regulatory quality and government effectiveness

```{r}
# install.packages("gganimate")
library(gganimate)

animation_selection <- esg_data_final %>% filter(
                          year >1999)

animation1 <- ggplot(data = animation_selection,
       mapping = aes(x = regulatory_quality_estimate,
                     y = government_effectiveness_estimate,
                     colour = continent,)) +
  geom_point(alpha = 0.5) +
  theme_bw() +
  theme(legend.position="none") +
  labs(title = "Year: {frame_time}", 
       x = "Internet Usage", 
       y = "Government Effectiveness Estimate") +    
  transition_time(year)+
  ease_aes("linear")

animate(animation1, height=600, width = 600, duration=20)
```

```{r}
countries_list <- c("China","United Kingdom","India","United States","South Africa", "Spain")

our_data <-esg_data_final %>% filter(country_name %in% countries_list)

```
# Visualizations

#### Note

All other visualizations not shown here were done in Tableau using the cleaned data in this Rmd file.

## Visualization for slide 3

```{r, cleaning gdp data, message=FALSE, warning=FALSE}
gdp <- read_csv(here::here("data","world_gdp.csv"))
gdp <- gdp %>% clean_names()
gdp <- pivot_longer(gdp,x1970_yr1970:x2020_yr2020,names_to = c("year","Year"),values_to = "amount",names_pattern = "x(.*)_yr(.*)")
gdp <- gdp %>% select(country_name,country_code,series,year,amount)

gdp_df <- gdp %>%
  pivot_wider(id_cols = c(country_name,country_code,year),names_from = series,values_from = amount,values_fn = max) %>% 
  clean_names() %>% 
  select(-c(country_name,country_code,year)) %>% 
  mutate_if(is.character,as.numeric)

gdp_index <- gdp %>%
  pivot_wider(id_cols = c(country_name,country_code,year),names_from = series,values_from = amount,values_fn = max) %>% 
  select(country_name,country_code,year)
gdp_index$year = as.Date(ISOdate(gdp_index$year, 1, 1),format = "%Y")

gdp_final <- cbind(gdp_index,gdp_df)

write_csv(gdp_final,here::here("data","gdp_final.csv"))

```

```{r, ploting gdp with world average, message=FALSE, warning=FALSE}
world_gdp <- gdp_final %>% 
  filter(year>"1990-01-01") %>% 
  select(country_name,year,gdp_per_capita_ppp_current_international, gdp_ppp_current_international) %>% 
  group_by(year) %>% 
  summarise(avg_gdp_capita = mean(gdp_per_capita_ppp_current_international,na.rm = TRUE),
            avg_gdp = mean(gdp_ppp_current_international,na.rm = TRUE)) 

comparison_gdp <- gdp_final %>% 
  filter(year>"1990-01-01",
         country_name %in% c("China","India","United States","United Kingdom","Spain","South Africa")) %>% 
  select(country_name,year,gdp_per_capita_ppp_current_international,gdp_ppp_current_international)

gdp_plot <-  ggplot()+
  geom_line(data=comparison_gdp,aes(x=year,y=gdp_per_capita_ppp_current_international,color=country_name),
            size=0.5,
            show.legend = FALSE)+
  geom_line(data=world_gdp,aes(x=year,y=avg_gdp_capita),size=1,linetype = "dashed",color = "gray70")+
    labs(title = "There has been stable economic growth within the six countries since 1990", 
       subtitle = "GDP per capita evolution compared to world average",
       x="Year",
       y="GDP per capita PPP in $",
       size=13)+
  scale_colour_manual(values = c("India" = "#D0BC52","China" = "#FF6C67","Spain" = "#36D6D9","South Africa"= "#3CD175","United Kingdom"= "#8CBDFF","United States"= "#FF90F1"))+
  geom_text(
    data = data.frame(x = rep(as.Date("2019-01-1"),6), y = c(9000,13000,19000,40000,46000,62000), label = c("India","South Africa","China","Spain","United Kingdom","United States")),
    aes(x = x, y = y, label = label, fontface = 4),
    colour=c("#D0BC52","#3CD175","#FF6C67","#36D6D9","#8CBDFF","#FF90F1"),
    family="Arial",
    hjust = 0,
    lineheight = .3,
    size = 3,
    inherit.aes = FALSE,
  ) +
  geom_text(
    aes(x = as.Date("2019-01-1"), y = 22000, label = "World",fontface=4),
    colour="gray70",
    family="Arial",
    hjust = 0,
    lineheight = .3,
    size = 4,
    inherit.aes = FALSE,
  ) +
    annotate("rect",xmin=as.Date("2008-1-1"),xmax=as.Date("2009-1-1"),ymin=-Inf, ymax=Inf, fill="gray",alpha=0.5)+
    geom_text(
    aes(x = as.Date("2008-06-1"), y = 60000, label = "2008 Financial Crisis",fontface=3),
    colour="black",
    family="Arial",
    hjust = 0.5,
    lineheight = .3,
    size = 3,
    inherit.aes = FALSE,
  ) +
  xlim(as.Date("1990-01-01"),as.Date("2023-01-01"))+
  theme(legend.position = "none",
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_line(colour = "grey"),
        panel.background=element_blank(),
        axis.line = element_line(colour = "grey"),
        plot.background=element_blank())
```

```{r, plotting CO2 data, message=FALSE, warning=FALSE}
world_co2 <- esg_data_final %>% 
  select(country_name,year, co2_emissions_metric_tons_per_capita) %>% 
  filter(year>1990,
         !is.na(co2_emissions_metric_tons_per_capita)) %>% 
  group_by(year) %>% 
  summarise(avg_emission = mean(co2_emissions_metric_tons_per_capita,na.rm=TRUE)) 
  
comparison_co2_five <- esg_data_final %>% 
  filter(year>1990,
         country_name %in% c("India","United States","United Kingdom","Spain","South Africa")) 

comparison_co2_china <- esg_data_final %>% 
  filter(year>1990,
         country_name =="China") 

co2_plot <-   ggplot()+
  geom_line(data=world_co2,aes(x=year,y=avg_emission),
            size=1,linetype = "dashed",color = "gray70")+
  geom_line(data=comparison_co2_five,
            aes(x=year,y=co2_emissions_metric_tons_per_capita,color=country_name),
            show.legend = FALSE,
            size=0.5,
            alpha=0.7)+
  scale_colour_manual(values = c("India" = "#D0BC52","Spain" = "#36D6D9","South Africa"= "#3CD175","United Kingdom"= "#8CBDFF","United States"= "#FF90F1"))+
  geom_line(data=comparison_co2_china,
            aes(x=year,y=co2_emissions_metric_tons_per_capita),color="#FF6C67",size=1)+
      labs(title="China's rapid economic growth is partly at the expense of the environment",
       subtitle = "CO2 emission per capita evolution compared to world average",
       x="Year",
       y="CO2 emission per capita in T",
       size=13)+
  geom_text(
    data = data.frame(x = rep(1989,6), y = c(0.7,2.3,6,7,10.5,20), 
                      label = c("India","China","Spain","South Africa","United Kingdom","United States")),
    aes(x = x, y = y, label = label, fontface = 4),
    colour=c("#D0BC52","#FF6C67","#36D6D9","#3CD175","#8CBDFF","#FF90F1"),
    family="Arial",
    hjust = 0,
    lineheight = .3,
    size = 3,
    inherit.aes = FALSE,
  ) +
    geom_text(
    aes(x = 2019, y = 4, label = "World",fontface=4),
    colour="gray70",
    family="Arial",
    hjust = 0,
    lineheight = .3,
    size = 4,
    inherit.aes = FALSE,
  ) +
  theme(legend.position = "none",
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_line(colour = "grey"),
        panel.background=element_blank(),
        axis.line = element_line(colour = "grey"),
        plot.background=element_blank())

co2_plot
grid.arrange(gdp_plot,co2_plot,nrow=2)

```

## Visualization for slide 4

```{r, renewable energy output, message=FALSE, warning=FALSE}

renewable <- esg_data_final %>% 
  select(country_name,year,renewable_electricity_output_percent_of_total_electricity_output) %>% 
  filter(country_name %in% c("China","India","United States","United Kingdom","Spain","South Africa"),
         !is.na(renewable_electricity_output_percent_of_total_electricity_output)) %>% 
  mutate(UK = ifelse(country_name == "United Kingdom", TRUE, FALSE)) %>% 
  ggplot(aes(x=year,y=renewable_electricity_output_percent_of_total_electricity_output,color=country_name)) +
  geom_line(
    data = function(x) filter(x, UK),
    size = 1,
    color = "seagreen"
  ) +
  geom_line(data = function(x) filter(x, !UK),
            size=0.5,alpha = 0.7,show.legend = FALSE)+
  scale_colour_manual(values = c("India" = "#D0BC52","China" ="#FF6C67", "Spain" = "#36D6D9","South Africa"= "#3CD175","United States"= "#FF90F1"))+
  labs(title = "Renewable electricity output in UK has seen a rapid increase since 2010",
       subtitle = "Renewable electricity output as percentage of total electricity output",
       x="Year",
       y="Renewable electricity percentage",
       size=13)+
  theme(legend.position = "none",
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_line(colour = "grey"),
        panel.background=element_blank(),
        axis.line = element_line(colour = "grey"),
        plot.background=element_blank())+
    annotate("rect",xmin=2010,xmax=2015,ymin=-Inf, ymax=Inf, fill="palegreen4",alpha=0.3)+
  geom_text(
    data = data.frame(x = rep(1989,6), y = c(-1,4,12,18,22,26), label = c("South Africa","United Kingdom","United States","Spain","China","India")),
    aes(x = x, y = y, label = label, fontface=4),
    colour=c("#3CD175", "seagreen", "#FF90F1","#36D6D9", "#FF6C67", "#D0BC52"),
    family="Arial",
    hjust = 0,
    lineheight = .3,
    size = 3,
    inherit.aes = FALSE,
  )

renewable
```



```{r, UK greenhouse gas contribution, message=FALSE, warning=FALSE}
greenhouse <- readxl::read_xls(here::here("data","Greenhouse_UK.xls"))
greenhouse <- greenhouse %>% clean_names() %>% 
  mutate(
         manufacture = as.numeric(manufacturing),
         electricity_AC_supply = as.numeric(electricity_gas_steam_and_air_conditioning_supply),
         transport_storage = as.numeric(transport_and_storage),
         households = as.numeric(households)) %>% 
  select(year,manufacture,electricity_AC_supply,transport_storage,households)

greenhouse$year = as.Date(ISOdate(greenhouse$year, 1, 1),format = "%Y")

greenhouse <- greenhouse %>% pivot_longer(manufacture:households,names_to = "industry",values_to = "emission")

plot_uk <- ggplot(greenhouse,aes(x=year,y=emission,fill=factor(industry, levels = c("manufacture","electricity_AC_supply","households","transport_storage"))))+
  geom_area(alpha = 0.8,show.legend = FALSE)+
  scale_fill_manual("Industry",values=c("darkseagreen4","darkseagreen3","lightgoldenrod3","wheat4"))+
  labs(title = str_wrap("Energy Supply sector continues its downward trend in contribution to overall UK greenhouse gas emission",85),
       subtitle = "UK Greenhouse gas emissions for the three highest-emitting industries, and households",
       x="Year",
       y="Greenhouse gas emission (MT CO2)",
       size=13)+
  geom_text(
    data = data.frame(x = rep(as.Date("1995-01-1"),4), y = c(50,175,320,500), label = c("Manufacturing","Households","Energy Supply","Transport and Storage")),
    aes(x = x, y = y, label = label, fontface=4),
    color= c("gray40","gray40","white","gray40"),
    family="Arial",
    hjust = 0,
    lineheight = .5,
    size = 3,
    inherit.aes = FALSE,
  ) +
    theme(legend.position = "none",
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_line(colour = "grey"),
        panel.background=element_blank(),
        axis.line = element_line(colour = "grey"),
        plot.background=element_blank())

grid.arrange(renewable,plot_uk,nrow=2)


```

## Visualization of correlation for government effectiveness and regulatory quality

```{r}
library(GGally)
esg_data_final %>% 
  select(country_name, year,ease_of_doing_business_rank_1_most_business_friendly_regulations,
         government_effectiveness_estimate, regulatory_quality_estimate) %>% #keep Y variable last
  ggcorr(method = c("pairwise", "pearson"), layout.exp = 3,label_round=2, label = TRUE,label_size = 2,hjust = 1)


esg_data_final %>% filter(continent != "NA") %>% ggplot(aes(x=regulatory_quality_estimate, y=government_effectiveness_estimate, color=continent)) +
  geom_point() + labs(
       title = "Almost perfect correlation between Government Effectiveness and Regulatory Quality",
       subtitle = "Scatterplot of yearly values for each country grouped by continent",
       caption = "Source: World Bank ESG Data",
       x= "Regulatory Quality Estimate",
       y = "Government Effectiveness Estimate",
       color="Continent") +
  
  theme(panel.background=element_blank(),
          panel.border=element_blank(),
          plot.background=element_blank()) +
  theme(axis.line = element_line(colour = "black")) +
    theme(text=element_text(size=11, family="Arial")) +
    theme(plot.title.position = "plot") +
    theme(plot.subtitle = element_text( size=11)) +
  geom_text(
    data = data.frame(x = -1, y = 1.5, label = "2019"),
    aes(x = x, y = y, label = "Correlation = 0.93"),
    colour="black",
    family="Arial",
    hjust = 0.5,
    lineheight = .8,
    inherit.aes = FALSE,
  )


```
### Comparison over the years

```{r}
#regulatory quality average
world_avg_reg<- esg_data_final %>% 
  select(country_name, regulatory_quality_estimate, year, continent) %>%
  filter(regulatory_quality_estimate != "NA", continent != "NA") %>%
  group_by(year) %>%
  summarize(world_avg = mean(regulatory_quality_estimate),
            country_name = "World")

world_avg_govteffec<- esg_data_final %>% 
  select(country_name, government_effectiveness_estimate, year, continent) %>%
  filter(government_effectiveness_estimate != "NA", continent != "NA") %>%
  group_by(year) %>%
  summarize(world_avg = mean(government_effectiveness_estimate),
            country_name = "World")


one<-our_data %>% filter(government_effectiveness_estimate != "NA") %>% ggplot(aes(y=government_effectiveness_estimate, x=year, group=country_name, color=country_name)) + 
  geom_smooth(se=FALSE) +
  geom_smooth(data=world_avg_reg, aes(x=year, y=world_avg), linetype = "dashed",color = "gray70") + labs(
       title = "Comparison of Government Effectiveness and Regulatory Quality",
       subtitle = "Lineplots over years for selected countries compared to world average",
       x= "",
       y = "Government Effectiveness Estimate",
       color="Country") +
  theme(panel.background=element_blank(),
          panel.border=element_blank(),
          plot.background=element_blank()) +
  theme(axis.line = element_line(colour = "black")) +
    theme(text=element_text(size=14, family="Arial")) +
    theme(plot.title.position = "plot",
          legend.position = "none") +
    theme(plot.subtitle = element_text( size=11),
           axis.title=element_text(size=12)) +
  scale_y_continuous(breaks=c(0,1,2)) +
   geom_text(
    data = data.frame(x = c(2019,2012,2018.5,2018,2016,2013), y = c(.5,.5,.75,1.15,1.7,1.4), label = c("India","South Africa","China","Spain","United Kingdom","United States")),
    aes(x = x, y = y, label = label, fontface=4),
    colour=c("#D0BC52","#3CD175","#FF6C67","#36D6D9","#8CBDFF","#FF90F1"),
    family="Arial",
    hjust = 0,
    lineheight = .3,
    size = 3,
    inherit.aes = FALSE,
  ) +
    geom_text(
    aes(x = 2018, y = 0, label = "World",fontface=4),
    colour="gray70",
    family="Arial",
    hjust = 0,
    lineheight = .3,
    size = 4,
    inherit.aes = FALSE,
  )





 two<- our_data %>% filter(regulatory_quality_estimate != "NA") %>% ggplot(aes(y=regulatory_quality_estimate, x=year, group=country_name, color=country_name)) + 
  geom_smooth(se=FALSE) + 
   geom_smooth(data=world_avg_reg, aes(x=year, y=world_avg), linetype = "dashed",color = "gray70") +
  labs(
       title = "",
       x= "",
       y = "Regulatory Quality Estimate",
       color="Country") +
  theme(panel.background=element_blank(),
          panel.border=element_blank(),
          plot.background=element_blank()) +
  theme(axis.line = element_line(colour = "black")) +
    theme(text=element_text(size=14, family="Arial")) +
    theme(plot.title.position = "plot",
          legend.position = "none") +
    theme(plot.subtitle = element_text( size=11),
          axis.title=element_text(size=12)) +
   geom_text(
    data = data.frame(x = c(2018.5,2016,2018.5,2018,2017,2016), y = c(-.5,.4,-.4,1,1.8,1.5), label = c("India","South Africa","China","Spain","United Kingdom","United States")),
    aes(x = x, y = y, label = label, fontface=4),
    colour=c("#D0BC52","#3CD175","#FF6C67","#36D6D9","#8CBDFF","#FF90F1"),
    family="Arial",
    hjust = 0,
    lineheight = .3,
    size = 3,
    inherit.aes = FALSE,
  ) +
    geom_text(
    aes(x = 2018, y = 0, label = "World",fontface=4),
    colour="gray70",
    family="Arial",
    hjust = 0,
    lineheight = .3,
    size = 4,
    inherit.aes = FALSE,
  )
   
 
one
two

one/two
#put spain
#change style
#china election and south africa explanation
```

## Visualizations for Migration & Political Stability

```{r, clean governance data, echo=TRUE}
gov_data_final <- esg_data_final %>% select(country_name, country_code, year, net_migration, political_stability_and_absence_of_violence_terrorism_estimate) %>% 
  mutate(net_migration = net_migration/1000000) %>% 
  filter(country_name %in% c("China","India","United States","United Kingdom","Spain","South Africa")) %>% 
  filter ( !is.na (net_migration)) %>% 
  filter( !is.na( political_stability_and_absence_of_violence_terrorism_estimate))

```


``` {r, echo=TRUE}
p1 <- gov_data_final  %>% ggplot(aes(x=year,y=net_migration,color=country_name) , format(y, scientific = FALSE))+
  geom_line()+
  #facet_wrap(~country_name) +
  theme(legend.position = "bottom") +
  theme_minimal()

p1

```

```{r}


#Calculating the AVERAGE mig for the countries

gov_averages <- gov_data_final %>% 
              group_by(country_name) %>%   
                mutate ( total_mig = sum(net_migration) ) %>% 
               mutate ( total_pol = sum(political_stability_and_absence_of_violence_terrorism_estimate) ) %>% 
      #diving by 20 to average the migration across 10 years since data is from 2002 - 2012
              mutate (avg_across_migration = total_mig / 10) %>% 
              mutate (avg_across_political_stability = total_pol / 10) %>% 
              group_by(country_name)

gov_averages <- gov_averages %>% 
                  filter( year %in% c("2017" ) ) %>% 
                  select ( country_name , year, avg_across_migration, avg_across_political_stability )


```




```{r include=FALSE}
ggplot(gov_averages, x= country_name, aes(avg_across_migration , avg_across_political_stability, total, fill = country_name)) + 
geom_bar(stat="identity", position="dodge", color="black") 

```


```{r}


gov_averages %>% 
  ggplot(aes(y = avg_across_migration, x = fct_reorder(country_name, -avg_across_migration))) +
 #geom_line( aes( x = fct_reorder(country_name, -avg_across_migration) , y = avg_across_political_stability)) +
  scale_fill_brewer(palette = "Blues")+
  geom_col() +
  labs(
    title = "Are people migrating to the most politically stable countries?",
    subtitle = "Political stability and migration rates per country",
    x = "Country",
    y = "Migration & Stability"
  ) +  theme_minimal() 

```






```{r}


migration_plot <-  ggplot()+
  geom_line(data=gov_data_final,aes(x=year,y=net_migration,color=country_name),
            size=0.5,
            show.legend = TRUE)  +
    labs(title = "Highest migration seen in the western countries", 
       subtitle = "Net migration in countries over time",
       x="Year",
       y="Migration Per Million",
       size=13)+
  scale_colour_manual(values = c("India" = "#D0BC52","China" = "#FF6C67","Spain" = "#36D6D9","South Africa"= "#3CD175","United Kingdom"= "#8CBDFF","United States"= "#FF90F1"))+
   theme(legend.position = "bottom",
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.background=element_blank(),
        axis.line = element_line(colour = "grey"),
        plot.background=element_blank()) + scale_fill_manual("legend_title")
  NULL



stability_plot <-  ggplot()+
  geom_line(data=gov_data_final,aes(x=year,y=political_stability_and_absence_of_violence_terrorism_estimate
,color=country_name),
            size=0.5,
            show.legend = TRUE)  +
    labs(title = "US, UK & Spain have been stable over years", 
       subtitle = "Net stability in countries over time",
       x="Year",
       y="Political Stability",
       size=13)+
  scale_colour_manual(values = c("India" = "#D0BC52","China" = "#FF6C67","Spain" = "#36D6D9","South Africa"= "#3CD175","United Kingdom"= "#8CBDFF","United States"= "#FF90F1"))+
   theme(legend.position = "bottom",
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.background=element_blank(),
        axis.line = element_line(colour = "grey"),
        plot.background=element_blank()) +
  NULL

grid.arrange(migration_plot,stability_plot,ncol=2)

```




